What's in a name? On the automated topic naming of software maintenance activities

### 2010-03-06 icpc/icpc2010.tex
hi-res version..
easterbrook format for abs

\begin{abstract}
  Within the field of mining software repositories, approaches such as topic modeling and concept location rely on the automated   application of machine learning algorithms to corpora.  %What is a 'topic'?

% Other approaches

This kind of approach typically makes no use of domain-specific knowledge to improve the results or put them into a context that a developer can relate to.  While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail.

 In this paper, we investigate what topics the latent topic extractors produce, and examine whether such topics can be generalized across software products.  We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  In our approach, we name topics based on a static ontology of non-functional requirements and software qualities; in essence, this adds context to general-purpose ontologies such as WordNet.  In this paper, we discuss how we use this software development domain knowledge to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS.  Since our taxonomy is global in scope, our results allow us to compare the relative importance of particular qualities across different projects. Within the topics extracted from MySQL and MaxDB we found that non-functional requirements were common and often were trends across time. Our automated techniques can label topics without much computational overhead.

@@

% Clean this up
Traditionally topic extraction has required manual annotation to derive domain-relevant labels. This paper attempts to implement \emph{labelled topic extraction}, in which topic are extracted and given labels relating to a cross-project taxonomy. This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics. We also compare this approach to machine learners.
%We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology. From the annotated topics, we %apply automated classifiers to find labels for the remaining topics.

@@

Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
%There is some debate about the significance and importance of the %terms in this model. 
%However, it is ``an international standard and thus provides an %internationally accepted terminology for software quality \cite[p. %58]{Boegh2008},'' which is sufficient for the purposes of this %research. 
ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.
However, the terms extracted from ISO9126 may not capture all words associated with the labels.  For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. We therefore took the words from the taxonomy and expanded them.

>>>>

\begin{abstract}
To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for work topics. We can then categorize development artifacts as belonging to one or more of these topics. 
% Other approaches
In order to do this, approaches such as topic modeling and concept location rely on the automated application of machine learning algorithms to development artifacts. 
%limitations
However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.  While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
%introduce idea
This paper implements \emph{labelled topic extraction}, in which topics are extracted and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  In our approach, we name topics based on a static ontology of non-functional requirements and software qualities; in essence, this adds context to general-purpose ontologies such as WordNet.  
%elaborate
In this paper, we discuss how we use this software development domain knowledge to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS.  Since our taxonomy is global in scope, our results allow us to compare the relative importance of particular qualities across different projects. 
%sum up
Within the topics extracted from MySQL and MaxDB we found that non-functional requirements were common and often were trends across time. Our automated techniques can label topics with little computational overhead.

@@

% Clean this up
Traditionally topic extraction has required manual annotation to derive domain-relevant labels. This paper implements \emph{labelled topic extraction}, in which topics are extracted and given labels relating to a cross-project taxonomy. This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics. We also compare this approach to machine learners.
%We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology. From the annotated topics, we %apply automated classifiers to find labels for the remaining topics.

@@

Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
%There is some debate about the significance and importance of the %terms in this model. 
%However, it is ``an international standard and thus provides an %internationally accepted terminology for software quality \cite[p. %58]{Boegh2008},'' which is sufficient for the purposes of this %research. 
ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Bøegh2008},'' that is sufficient for the purposes of this research.
However, the terms extracted from ISO9126 may not capture all words associated with the labels.  For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. We therefore took the words from the taxonomy and expanded them.
